---
title: "Result"
shorttitle        : "Effort Result"
author: 
  - name          : "Amber X. Chen"
    affiliation   : "1"
authornote: |
  Add complete correspondence information here.

abstract: |
  This is an abstract.

keywords          : "keywords, go, here"
wordcount         : "X"

figsintext        : no
fig_caption       : yes

documentclass     : "apa7"
classoption       : "man"
output: 
  papaja::apa6_docx:
    latex_engine: "xelatex"
---
---

## Arima

```{r}
mainpath <- "/home/local/PSYCH-ADS/xuqian_chen/Github/effort"
datafolder <- file.path(mainpath, "2_pipeline/out")
arimafolder <-  file.path(mainpath, "3_output/results/arima")
grangerfolder <-  file.path(mainpath,"3_output/results/granger")
library(forecast)
library(tseries)
# Granger test
library(lmtest)
library(zoo)
library(stargazer)
# VAR
library(vars)
library(papaja)

```

```{r}

dfus <- read.csv(file.path(datafolder,"merged_us.csv"))
dfchi <- read.csv(file.path(datafolder,"merged_chi.csv"))

# transform GPD per capita to log scale
dfus$gdp_per_capita_log <- log(dfus$GDP.per.capita)
dfchi$gdp_per_capita_log <- log(dfchi$GDP.per.capita)
dfus$gdp_per_capita_z <- scale(dfus$gdp_per_capita_log)
dfchi$gdp_per_capita_z <- scale(dfchi$gdp_per_capita_log)
# loop through dependent variables to run ARIMA, granger causality, and plot
dfus$effort <- dfus$effort_virtue - dfus$effort_vice
dfus$efficiency <- dfus$efficiency_virtue - dfus$efficiency_vice
dfus$InefficientEffort <- dfus$effort_virtue - dfus$efficiency_virtue
dfchi$effort <- dfchi$effort_virtue - dfchi$effort_vice
dfchi$efficiency <- dfchi$efficiency_virtue - dfchi$efficiency_vice
dfchi$InefficientEffort <- dfchi$effort_virtue - dfchi$efficiency_virtue


dfus$indi_z <- as.vector(scale(dfus$indi))
dfchi$indi_z <-  as.vector(scale(dfchi$indi))
dfus$coll_z <-  as.vector(scale(dfus$coll))
dfchi$coll_z <-  as.vector(scale(dfchi$coll))
dfus$loose_z <-  as.vector(scale(dfus$loose))
dfchi$loose_z <-  as.vector(scale(dfchi$loose))
dfus$tight_z <-  as.vector(scale(dfus$tight))
dfchi$tight_z <-  as.vector(scale(dfchi$tight))

gdp_var <- "gdp_per_capita_z"
```

```{r}
# Define the function to perform ARIMA modeling

perform_arima_stats <- function(data, dv, iv, nouseivname,gdp_var = NULL) {
  x_ts <- ts(data[iv])
  y_ts <- ts(data[dv])
  
  if (!is.null(gdp_var)) {
    gdp_ts <- ts(data[[gdp_var]])
    xreg <- cbind(gdp_ts, x_ts)
    fit <- auto.arima(y_ts, xreg = xreg)
    ar <- fit$arma[1]
    ma <- fit$arma[2]
    diff <- fit$arma[6]
    model <- arima(y_ts, order = c(ar, diff, ma), xreg = xreg)
    coef_val <- coef(model)['x_ts']
    se_val <- sqrt(diag(vcov(model)))['x_ts']

  } else {
    xreg <- cbind(x_ts)
    fit <- auto.arima(y_ts, xreg = xreg)
    ar <- fit$arma[1]
    ma <- fit$arma[2]
    diff <- fit$arma[6]
    model <- arima(y_ts, order = c(ar, diff, ma), xreg = xreg)
    coef_val <- coef(model)[iv]
    se_val <- sqrt(diag(vcov(model)))[iv]

  }

  ci_low <- coef_val - 1.96 * se_val
  ci_high <- coef_val + 1.96 * se_val
  p_value <- 2 * (1 - pnorm(abs(coef_val/se_val)))
  return(list(coef = coef_val, ci_low = ci_low, ci_high = ci_high, p_value = p_value))
}
# Extracting and formatting values for clarity:
format_stats <- function(model) {
    coef <- sprintf("%.2f", model$coef)
    ci_low <- sprintf("%.2f", model$ci_low)
    ci_high <- sprintf("%.2f", model$ci_high)
    p_value <- ifelse(model$p_value < 0.001, "< 0.001", sprintf("%.3f", model$p_value))
    return(paste("(*b* = ", coef, ", 95% CI [", ci_low, ", ", ci_high, "], *p* = ", p_value, ")", sep = ""))
}
# Define the function to perform Granger causality test and get formatted statistics
perform_granger_stats <- function(data, iv, dv) {
  # Combine the series into a data frame
  df <- data.frame(y = data[dv], x = data[iv])
  
  # Remove rows with NA values
  df <- df[complete.cases(df), ]
  
  # Determine the best order for the VAR model using AIC
  var_result <- VARselect(df, lag.max = 10, type = "both")
  optimal_lag <- var_result$selection["AIC(n)"]
  
  # Fit the VAR model
  var_model <- VAR(df, p = optimal_lag)
  
  # Perform Granger causality test using the estimated VAR model
  granger_test <- causality(var_model, cause = iv)
  
  # Extract statistics
  p_value <- granger_test$Granger$p.value
  f_value <- granger_test$Granger$statistic[[1]]
  df1 <- granger_test$Granger$parameter[[1]]
  df2 <- granger_test$Granger$parameter[[2]]
  
  # Format for easy inclusion in text
  p_str <- ifelse(p_value < 0.001, "< 0.001", sprintf("%.3f", p_value))
  formatted_stats <- sprintf("(*F*(%d, %d) = %.2f, *p* = %s)", df1, df2, f_value, p_str)

  # Return results
  result <- list(optimal_lag = optimal_lag, formatted_stats = formatted_stats, p_value = p_value)
  return(result)
}

```
```{r}
useffortindi <- perform_arima_stats(dfus,"effort","indi_z","Individualism",gdp_var)
useffortcoll <- perform_arima_stats(dfus,"effort","coll_z","Collectivism",gdp_var)
useffortloose <- perform_arima_stats(dfus,"effort","loose_z","Looseness",gdp_var)
usefforttight <- perform_arima_stats(dfus,"effort","tight_z","Tightness",gdp_var)
usefficiencyindi <- perform_arima_stats(dfus,"efficiency","indi_z","Individualism",gdp_var)
usefficiencycoll <- perform_arima_stats(dfus,"efficiency","coll_z","Collectivism",gdp_var)
usefficiencyloose <- perform_arima_stats(dfus,"efficiency","loose_z","Looseness",gdp_var)
usefficiencytight <- perform_arima_stats(dfus,"efficiency","tight_z","Tightness",gdp_var)
usInefficientEffortindi <- perform_arima_stats(dfus,"InefficientEffort","indi_z","Individualism",gdp_var)
usInefficientEffortcoll <- perform_arima_stats(dfus,"InefficientEffort","coll_z","Collectivism",gdp_var)
usInefficientEffortloose <- perform_arima_stats(dfus,"InefficientEffort","loose_z","Looseness",gdp_var)
usInefficientEfforttight <- perform_arima_stats(dfus,"InefficientEffort","tight_z","Tightness",gdp_var)

useffortindi_noGDP <- perform_arima_stats(dfus,"effort","indi_z","Individualism")
useffortcoll_noGDP <- perform_arima_stats(dfus,"effort","coll_z","Collectivism")
useffortloose_noGDP <- perform_arima_stats(dfus,"effort","loose_z","Looseness")
usefforttight_noGDP <- perform_arima_stats(dfus,"effort","tight_z","Tightness")
usefficiencyindi_noGDP <- perform_arima_stats(dfus,"efficiency","indi_z","Individualism")
usefficiencycoll_noGDP <- perform_arima_stats(dfus,"efficiency","coll_z","Collectivism")
usefficiencyloose_noGDP <- perform_arima_stats(dfus,"efficiency","loose_z","Looseness")
usefficiencytight_noGDP <- perform_arima_stats(dfus,"efficiency","tight_z","Tightness")
usInefficientEffortindi_noGDP <- perform_arima_stats(dfus,"InefficientEffort","indi_z","Individualism")
usInefficientEffortcoll_noGDP <- perform_arima_stats(dfus,"InefficientEffort","coll_z","Collectivism")
usInefficientEffortloose_noGDP <- perform_arima_stats(dfus,"InefficientEffort","loose_z","Looseness")
usInefficientEfforttight_noGDP <- perform_arima_stats(dfus,"InefficientEffort","tight_z","Tightness")


chieffortindi <- perform_arima_stats(dfchi,"effort","indi_z","Individualism",gdp_var)
chieffortcoll <- perform_arima_stats(dfchi,"effort","coll_z","Collectivism",gdp_var)
chieffortloose <- perform_arima_stats(dfchi,"effort","loose_z","Looseness",gdp_var)
chiefforttight <- perform_arima_stats(dfchi,"effort","tight_z","Tightness",gdp_var)
chiefficiencyindi <- perform_arima_stats(dfchi,"efficiency","indi_z","Individualism",gdp_var)
chiefficiencycoll <- perform_arima_stats(dfchi,"efficiency","coll_z","Collectivism",gdp_var)
chiefficiencyloose <- perform_arima_stats(dfchi,"efficiency","loose_z","Looseness",gdp_var)
chiefficiencytight <- perform_arima_stats(dfchi,"efficiency","tight_z","Tightness",gdp_var)
chiInefficientEffortindi <- perform_arima_stats(dfchi,"InefficientEffort","indi_z","Individualism",gdp_var)
chiInefficientEffortcoll <- perform_arima_stats(dfchi,"InefficientEffort","coll_z","Collectivism",gdp_var)
chiInefficientEffortloose <- perform_arima_stats(dfchi,"InefficientEffort","loose_z","Looseness",gdp_var)
chiInefficientEfforttight <- perform_arima_stats(dfchi,"InefficientEffort","tight_z","Tightness",gdp_var)

chieffortindi_noGDP <- perform_arima_stats(dfchi,"effort","indi_z","Individualism")
chieffortcoll_noGDP <- perform_arima_stats(dfchi,"effort","coll_z","Collectivism")
chieffortloose_noGDP <- perform_arima_stats(dfchi,"effort","loose_z","Looseness")
chiefforttight_noGDP <- perform_arima_stats(dfchi,"effort","tight_z","Tightness")
chiefficiencyindi_noGDP <- perform_arima_stats(dfchi,"efficiency","indi_z","Individualism")
chiefficiencycoll_noGDP <- perform_arima_stats(dfchi,"efficiency","coll_z","Collectivism")
chiefficiencyloose_noGDP <- perform_arima_stats(dfchi,"efficiency","loose_z","Looseness")
chiefficiencytight_noGDP <- perform_arima_stats(dfchi,"efficiency","tight_z","Tightness")
chiInefficientEffortindi_noGDP <- perform_arima_stats(dfchi,"InefficientEffort","indi_z","Individualism")
chiInefficientEffortcoll_noGDP <- perform_arima_stats(dfchi,"InefficientEffort","coll_z","Collectivism")
chiInefficientEffortloose_noGDP <- perform_arima_stats(dfchi,"InefficientEffort","loose_z","Looseness")
chiInefficientEfforttight_noGDP <- perform_arima_stats(dfchi,"InefficientEffort","tight_z","Tightness")

# granger
grangeruseffortindi <- perform_granger_stats(dfus,"effort","indi_z")
grangerusindieffort <- perform_granger_stats(dfus,"indi_z","effort")
grangeruseffortcoll <- perform_granger_stats(dfus,"effort","coll_z")
grangeruscolleffort <- perform_granger_stats(dfus,"coll_z","effort")
grangeruseffortloose <- perform_granger_stats(dfus,"effort","loose_z")
grangeruslooseeffort <- perform_granger_stats(dfus,"loose_z","effort")
grangerusefforttight <- perform_granger_stats(dfus,"effort","tight_z")
grangerustighteffort <- perform_granger_stats(dfus,"tight_z","effort")
grangerusefficiencyindi <- perform_granger_stats(dfus,"efficiency","indi_z")
grangerusindiefficiency <- perform_granger_stats(dfus,"indi_z","efficiency")
grangerusefficiencycoll <- perform_granger_stats(dfus,"efficiency","coll_z")
grangeruscollefficiency <- perform_granger_stats(dfus,"coll_z","efficiency")
grangerusefficiencyloose <- perform_granger_stats(dfus,"efficiency","loose_z")
grangeruslooseefficiency <- perform_granger_stats(dfus,"loose_z","efficiency")
grangerusefficiencytight <- perform_granger_stats(dfus,"efficiency","tight_z")
grangerustightefficiency <- perform_granger_stats(dfus,"tight_z","efficiency")
grangerusInefficientEffortindi <- perform_granger_stats(dfus,"InefficientEffort","indi_z")
grangerusindiInefficientEffort <- perform_granger_stats(dfus,"indi_z","InefficientEffort")
grangerusInefficientEffortcoll <- perform_granger_stats(dfus,"InefficientEffort","coll_z")
grangeruscollInefficientEffort <- perform_granger_stats(dfus,"coll_z","InefficientEffort")
grangerusInefficientEffortloose <- perform_granger_stats(dfus,"InefficientEffort","loose_z")
grangeruslooseInefficientEffort <- perform_granger_stats(dfus,"loose_z","InefficientEffort")
grangerusInefficientEfforttight <- perform_granger_stats(dfus,"InefficientEffort","tight_z")
grangerustightInefficientEffort <- perform_granger_stats(dfus,"tight_z","InefficientEffort")

grangerchieffortindi <- perform_granger_stats(dfchi,"effort","indi_z")
grangerchiindieffort <- perform_granger_stats(dfchi,"indi_z","effort")
grangerchieffortcoll <- perform_granger_stats(dfchi,"effort","coll_z")
grangerchicolleffort <- perform_granger_stats(dfchi,"coll_z","effort")
grangerchieffortloose <- perform_granger_stats(dfchi,"effort","loose_z")
grangerchilooseeffort <- perform_granger_stats(dfchi,"loose_z","effort")
grangerchiefforttight <- perform_granger_stats(dfchi,"effort","tight_z")
grangerchitighteffort <- perform_granger_stats(dfchi,"tight_z","effort")
grangerchiefficiencyindi <- perform_granger_stats(dfchi,"efficiency","indi_z")
grangerchiindiefficiency <- perform_granger_stats(dfchi,"indi_z","efficiency")
grangerchiefficiencycoll <- perform_granger_stats(dfchi,"efficiency","coll_z")
grangerchicollefficiency <- perform_granger_stats(dfchi,"coll_z","efficiency")
grangerchiefficiencyloose <- perform_granger_stats(dfchi,"efficiency","loose_z")
grangerchilooseefficiency <- perform_granger_stats(dfchi,"loose_z","efficiency")
grangerchiefficiencytight <- perform_granger_stats(dfchi,"efficiency","tight_z")
grangerchitightefficiency <- perform_granger_stats(dfchi,"tight_z","efficiency")
grangerchiInefficientEffortindi <- perform_granger_stats(dfchi,"InefficientEffort","indi_z")
grangerchiindiInefficientEffort <- perform_granger_stats(dfchi,"indi_z","InefficientEffort")
grangerchiInefficientEffortcoll <- perform_granger_stats(dfchi,"InefficientEffort","coll_z")
grangerchicollInefficientEffort <- perform_granger_stats(dfchi,"coll_z","InefficientEffort")
grangerchiInefficientEffortloose <- perform_granger_stats(dfchi,"InefficientEffort","loose_z")
grangerchilooseInefficientEffort <- perform_granger_stats(dfchi,"loose_z","InefficientEffort")
grangerchiInefficientEfforttight <- perform_granger_stats(dfchi,"InefficientEffort","tight_z")
grangerchitightInefficientEffort <- perform_granger_stats(dfchi,"tight_z","InefficientEffort")

```

## ARIMA Models: US Congressional Speeches

- Individualism predicts effort without GDP: `r format_stats(useffortindi_noGDP)`, controlling for GDP:`r format_stats(useffortindi)`
- Collectivism predicts effort without GDP: `r format_stats(useffortcoll_noGDP)`, controlling for GDP:`r format_stats(useffortcoll)`
- Looseness predicts effort without GDP: `r format_stats(useffortloose_noGDP)`, controlling for GDP:`r format_stats(useffortloose)`
- Tightness predicts effort without GDP: `r format_stats(usefforttight_noGDP)`, controlling for GDP:`r format_stats(usefforttight)`

- Individualism predicts efficiency without GDP: `r format_stats(usefficiencyindi_noGDP)`, controlling for GDP:`r format_stats(usefficiencyindi)`
- Collectivism predicts efficiency without GDP: `r format_stats(usefficiencycoll_noGDP)`, controlling for GDP:`r format_stats(usefficiencycoll)`
- Looseness predicts efficiency without GDP: `r format_stats(usefficiencyloose_noGDP)`, controlling for GDP:`r format_stats(usefficiencyloose)`
- Tightness predicts efficiency without GDP: `r format_stats(usefficiencytight_noGDP)`, controlling for GDP:`r format_stats(usefficiencytight)`

- Individualism predicts inefficient effort without GDP: `r format_stats(usInefficientEffortindi_noGDP)`, controlling for GDP:`r format_stats(usInefficientEffortindi)`
- Collectivism predicts inefficient effort without GDP: `r format_stats(usInefficientEffortcoll_noGDP)`, controlling for GDP:`r format_stats(usInefficientEffortcoll)`
- Looseness predicts inefficient effort without GDP: `r format_stats(usInefficientEffortloose_noGDP)`, controlling for GDP:`r format_stats(usInefficientEffortloose)`
- Tightness predicts inefficient effort without GDP: `r format_stats(usInefficientEfforttight_noGDP)`, controlling for GDP:`r format_stats(usInefficientEfforttight)`

## ARIMA Models: People’s Daily of China

- Individualism predicts effort without GDP: `r format_stats(chieffortindi_noGDP)`, controlling for GDP:`r format_stats(chieffortindi)`
- Collectivism predicts effort without GDP: `r format_stats(chieffortcoll_noGDP)`, controlling for GDP:`r format_stats(chieffortcoll)`
- Looseness predicts effort without GDP: `r format_stats(chieffortloose_noGDP)`, controlling for GDP:`r format_stats(chieffortloose)`
- Tightness predicts effort without GDP: `r format_stats(chiefforttight_noGDP)`, controlling for GDP:`r format_stats(chiefforttight)`

- Individualism predicts efficiency without GDP: `r format_stats(chiefficiencyindi_noGDP)`, controlling for GDP:`r format_stats(chiefficiencyindi)`
- Collectivism predicts efficiency without GDP: `r format_stats(chiefficiencycoll_noGDP)`, controlling for GDP:`r format_stats(chiefficiencycoll)`
- Looseness predicts efficiency without GDP: `r format_stats(chiefficiencyloose_noGDP)`, controlling for GDP:`r format_stats(chiefficiencyloose)`
- Tightness predicts efficiency without GDP: `r format_stats(chiefficiencytight_noGDP)`, controlling for GDP:`r format_stats(chiefficiencytight)`

- Individualism predicts inefficient effort without GDP: `r format_stats(chiInefficientEffortindi_noGDP)`, controlling for GDP:`r format_stats(chiInefficientEffortindi)`
- Collectivism predicts inefficient effort without GDP: `r format_stats(chiInefficientEffortcoll_noGDP)`, controlling for GDP:`r format_stats(chiInefficientEffortcoll)`
- Looseness predicts inefficient effort without GDP: `r format_stats(chiInefficientEffortloose_noGDP)`, controlling for GDP:`r format_stats(chiInefficientEffortloose)`
- Tightness predicts inefficient effort without GDP: `r format_stats(chiInefficientEfforttight_noGDP)`, controlling for GDP:`r format_stats(chiInefficientEfforttight)`

## Granger Tests: US Congressional Speeches
- Individualism causes Effort: `r grangerusindieffort$formatted_stats` 
- Effort causes Individualism: `r grangeruseffortindi$formatted_stats` 
- Collectivism cuases Effort: `r grangeruscolleffort$formatted_stats` 
- Effort causes Collectivism `r grangeruseffortcoll$formatted_stats` 
- Looseness causes Effort: `r grangeruslooseeffort$formatted_stats`
- Effort causes Looseness: `r grangeruseffortloose$formatted_stats`
- Tightness causes Effort: `r grangerustighteffort$formatted_stats`
- Effort causes Tightness: `r grangerusefforttight$formatted_stats`

- Individualism causes Efficiency: `r grangerusindiefficiency$formatted_stats`
- Efficiency causes Individualism: `r grangerusefficiencyindi$formatted_stats`
- Collectivism cuases  Efficiency: `r grangeruscollefficiency$formatted_stats`
- Efficiency causes Collectivism `r grangerusefficiencycoll$formatted_stats`
- Looseness causes Efficiency: `r grangeruslooseefficiency$formatted_stats`
- Efficiency causes Looseness: `r grangerusefficiencyloose$formatted_stats`
- Tightness causes Efficiency: `r grangerustightefficiency$formatted_stats`

- Individualism causes Inefficient Effort: `r grangerusindiInefficientEffort$formatted_stats`
- Inefficient Effort causes Individualism: `r grangerusInefficientEffortindi$formatted_stats`
- Collectivism cuases  Inefficient Effort: `r grangeruscollInefficientEffort$formatted_stats`
- Inefficient Effort causes Collectivism `r grangerusInefficientEffortcoll$formatted_stats`
- Looseness causes Inefficient Effort: `r grangeruslooseInefficientEffort$formatted_stats`
- Inefficient Effort causes Looseness: `r grangerusInefficientEffortloose$formatted_stats`
- Tightness causes Inefficient Effort: `r grangerustightInefficientEffort$formatted_stats`

## Granger Tests: People’s Daily of China
- Individualism causes Effort: `r grangerchiindieffort$formatted_stats`
- Effort causes Individualism: `r grangerchieffortindi$formatted_stats`
- Collectivism cuases Effort: `r grangerchicolleffort$formatted_stats`
- Effort causes Collectivism `r grangerchieffortcoll$formatted_stats`
- Looseness causes Effort: `r grangerchilooseeffort$formatted_stats`
- Effort causes Looseness: `r grangerchieffortloose$formatted_stats`
- Tightness causes Effort: `r grangerchitighteffort$formatted_stats`
- Effort causes Tightness: `r grangerchiefforttight$formatted_stats`

- Individualism causes Efficiency: `r grangerchiindiefficiency$formatted_stats`
- Efficiency causes Individualism: `r grangerchiefficiencyindi$formatted_stats`
- Collectivism cuases  Efficiency: `r grangerchicollefficiency$formatted_stats`
- Efficiency causes Collectivism `r grangerchiefficiencycoll$formatted_stats`
- Looseness causes Efficiency: `r grangerchilooseefficiency$formatted_stats`
- Efficiency causes Looseness: `r grangerchiefficiencyloose$formatted_stats`
- Tightness causes Efficiency: `r grangerchitightefficiency$formatted_stats`

- Individualism causes Inefficient Effort: `r grangerchiindiInefficientEffort$formatted_stats`
- Inefficient Effort causes Individualism: `r grangerchiInefficientEffortindi$formatted_stats`
- Collectivism cuases  Inefficient Effort: `r grangerchicollInefficientEffort$formatted_stats`
- Inefficient Effort causes Collectivism `r grangerchiInefficientEffortcoll$formatted_stats`
- Looseness causes Inefficient Effort: `r grangerchilooseInefficientEffort$formatted_stats`
- Inefficient Effort causes Looseness: `r grangerchiInefficientEffortloose$formatted_stats`
- Tightness causes Inefficient Effort: `r grangerchitightInefficientEffort$formatted_stats`
